Complex Analysis of Askaryan Radiation: UHE-$\nu$ Identification and Reconstruction via the Hilbert Envelope of Observed Signals

### 2025-08-16 template.tex
added rho vs. SNR, contour plot needs work

\item A Monte Carlo data set was generated using NuRadioMC for UHE-$\nu$ interactions in a cylindrical ice volume with a depth of 0.65 km and a radius of 0.850 km.  The UHE-$\nu$ energies ranged from 10-100 PeV, with roughly equal numbers of UHE-$\nu$ in each 10 PeV bin.  The detector included a single string of 8 RF dipole antennas, a filtered passpand of [80,1000] MHz, and 256 samples per channel at a 1 GHz sampling rate.  The RF trigger responded when 3 of the 8 voltage traces exceeded $\pm 3$ times the $v_{\rm rms}$ of the thermal noise within 256 ns.  For each triggered UHE-$\nu$ interaction, the Hilbert envelope of the coherently summed waveform (CSW) was calculated and correlated against Eqs. \ref{eq:final}-\ref{eq:final2}.  A quick optimization over the whole data set yielded 0.15 GHz and 0.025 GHz for the $f_0$ and $\gamma$ parameters, respectively.  Each individual correlation was maximized by varying the just the $\sigma_t$ parameter, yielding an optimized analytical envelope for each event.  A thermal noise waveform reflective of the detector properties was also generated using NuRadioMC tools for each UHE-$\nu$ event.  The noise waveform was also correlated against the optimized analytical envelope.  The probability distributions for the correlation coefficients are shown in Fig. \ref{fig:fig3}.
\begin{figure}
\includegraphics[width=0.5\textwidth,trim=3.25cm 8.25cm 4.5cm 9.0cm,clip=true]{Aug15_plot1.pdf}
\caption{\label{fig:fig3} (Black circles) Noise distribution. (Gray dashed line) Fitting function to noise distribution.  (Solid black line) Exponential fit to tail of noise distribution.  (Solid gray line) UHE-$\nu$ signal distribution.  (Dashed black line) Optimized correlation coefficient threshold value.}
In Fig. \ref{fig:fig3}, the black circles represent the normalized histogram of the correlation coefficent between the optimized analytic envelope and the thermal noise waveform.  A function of the form $x^2 \exp(-0.5 x^2)$ was fit to the black circles on the interval [0.1,1.0], and is represented by the gray dashed line.  A function of the form $exp(-x)$ was fit to the noise distribution on the interval [0.28,1] to capture the limiting behavior, and is represented by the solid black line.  The solid gray line represents the signal distribution, which peaks at a correlation value of 0.96.  The vertical black dashed line represents a threshold value, 0.438.  For the simulated UHE-$\nu$ sample, 92\% of CSWs meet or exceed this correlation value when fit to an optimized analytic envelope.  Assuming the solid black line correctly describes the tail of the probability distribution function for noise correlation, a threshold of 0.438 is equivalent to 0.95 noise events per detector over a period of 5 years, at a noise trigger rate of 1 Hz.

Correlation vs SNR
\end{itemize}

>>>>

\item A Monte Carlo data set was generated using NuRadioMC for UHE-$\nu$ interactions in a cylindrical ice volume with a depth of 0.65 km and a radius of 0.850 km.  The UHE-$\nu$ energies ranged from 10-100 PeV, with roughly equal numbers of UHE-$\nu$ in each 10 PeV bin.  The detector had one string of 8 RF dipole antennas.  Each RF channel had a filtered, amplified passpand of [80,1000] MHz, with 256 samples per channel at a 1 GHz sampling rate.  The RF trigger responded when 3 of the 8 voltage traces exceeded $\pm 3$ times the $v_{\rm rms}$ of the thermal noise within 256 ns.  For each signal, the Hilbert envelope of the coherently summed waveform (CSW) was calculated and correlated against Eqs. \ref{eq:final}-\ref{eq:final2}.  A quick optimization over the whole data set yielded 0.15 GHz and 0.025 GHz for the $f_0$ and $\gamma$ parameters, respectively.  Each correlation was maximized by varying the just the $\sigma_t$ parameter, yielding an optimized analytical envelope for each event.  A thermal noise waveform reflective of the channel properties was also generated for each event.  The noise waveform was also correlated against the optimized analytical envelope.  Both correlation distributions are shown in Fig. \ref{fig:fig3}.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,trim=3.25cm 8.25cm 4.5cm 9.0cm,clip=true]{Aug15_plot1.pdf}
\caption{\label{fig:fig3} (Black circles) Noise distribution. (Gray dashed line) Fitting function to noise distribution.  (Solid black line) Exponential fit to tail of noise distribution.  (Solid gray line) UHE-$\nu$ signal distribution.  (Dashed black line) Optimized correlation coefficient threshold value.}
\end{figure}

In Fig. \ref{fig:fig3}, the black circles represent the normalized histogram of the correlation coefficent between the optimized analytic envelope and the thermal noise waveform.  Functions of the forms $x^2 \exp(-0.5 x^2)$ and $\exp(-x)$ were fit to the noise distribution on the intervals [0.1,1.0] and [0.28,1], respectively.  The former is represented by the gray dashed line, and the latter is represented by the solid black line.  The solid gray line represents the signal distribution, which peaks at a correlation value of 0.96.  The vertical black dashed line represents a threshold of 0.438.  For the simulated UHE-$\nu$, 92\% of correlations between CSWs and optimized analytic envelopes are greater than or equal to this threshold.  Assuming the solid black line describes the noise distribution tail, with a noise trigger rate of 1 Hz, the threshold is equivalent to 0.95 noise events every 5 years.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{Aug15_plot2.png}
\caption{\label{fig:fig3} Larger font without cutting it off, add noise data}
\end{figure}

\end{itemize}
